---
title: "Class Reference"
author: "Loretta Pulwer"
output:
  html_document:
    # theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```


library(tidyverse)
── Attaching core tidyverse packages ────────────────────── tidyverse 2.0.0 ──
✔ dplyr     1.1.4     ✔ readr     2.1.5
✔ forcats   1.0.0     ✔ stringr   1.5.1
✔ ggplot2   3.4.4     ✔ tibble    3.2.1
✔ lubridate 1.9.3     ✔ tidyr     1.3.1
✔ purrr     1.0.2     
── Conflicts ──────────────────────────────────────── tidyverse_conflicts() ──
✖ dplyr::filter() masks stats::filter()
✖ dplyr::lag()    masks stats::lag()
ℹ Use the conflicted package to force all conflicts to become errors

> #library loads packages. In this case, it's the tidyverse

> install.packages('tidyverse')
Error in install.packages : Updating loaded packages
> #this functon installs packages 

> maryland_winred_contributions <- read_rds("maryland_winred.rds")
Error in readRDS(con, refhook = refhook) : cannot open the connection
In addition: Warning message:
In readRDS(con, refhook = refhook) :
  cannot open file 'maryland_winred.rds': No such file or directory
> #I don't have the file on my computer. But in theory, R studio is supposed to read to import the data from the given file 

glimpse(maryland_winred_contributions)
# gives overview of the rows and columns 

head(maryland_winred_contributions)
#shows the first 6 rows 

maryland_winred_contributions |>
  group_by(date) |>
  summarise(
    count_contribs = n()
    
# groups by a certain variable

 arrange(desc(count_contribs))

#arranges rows 

maryland_winred_contributions |>
  group_by(memo_text, amount) |>
  summarise(
    count_contribs = n()
    
#multiple grouping 

maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount)
    
    
#mean and median 
maryland_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
    
#mean, median, min, mas 

summary()
#do math on numeric columns. Will provide median, quartiles, mean, mas, min. 

colnames()
#list of all the columns 

select()
#only show me this set of columns 

#example
primary_24 |>
  select(office_name, office_district, candidate_name, party, county_name, election_day, votes)
  
  

mutate()
#create a new column 

#example 1 
mutate(
  percent_election_day = election_day/votes
  
#example 2 (change existing data)
standardized_maryland_cities <- maryland_cities |>
  mutate(
    upper_city = str_to_upper(city)
  
primary_24 |>
  select(office_name, office_district, candidate_name, party, county_name, election_day, votes) |>
  mutate(
  percent_election_day = (election_day/votes)*100
# *100 moves the decimal   

str_to_upper
# puts everything into upper case "THE QUICK BROWN DOG"

str_to_lower
# puts everything into lower case "the quick brown dog"

str_to_title 
# capitalizes every first letter of a word 

case_when()
#sets conditions 

#example
maryland_winred_categories <- maryland_winred_contributions |>
  mutate(
    amount_category = case_when(
        amount < 100 ~ "Less than $100",
        amount >= 100 & amount < 500 ~ "Between $100 and $499",
        amount >= 500 & amount < 1500 ~ "Between $500 and $1499",
        amount >= 1500 & amount < 3000 ~ "Between $500 and $2999",
        amount >= 3000 ~ "$3,000 or more"
 
 filter()
#filters out certain info that is wanted 
 
#example
 filter(county_name == "Prince George's County")
 
#example 2 (combining filters)
 filter(county_name == "Prince George's County") |> filter(office_name == "U.S. Senator")
 
 
 AND filter
#example
 filter(county_name == "Prince George's County" & office_name == "U.S. Senator")
 
 OR filter 
#example 
 filter(county_name == "Prince George's County" | office_name == "U.S. Senator")
 
 
 mdy
#changes dates into YYYY-MM-DD format
#example
 -> mutate(expenditure_date=mdy(expenditure_date))
 
 maryland_expenses |>
  mutate(month = floor_date(expenditure_date, "month")) |>
  group_by(month) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))d
  
  read_cvs(guess_max=10)
#tells what kind of datatypes we have for the first 10 columns.
  
  texas_precinct_20 <- texas_precinct_20 |>
  mutate(election_day = case_when(
    election_day == '5+' ~ '49',
    .default = election_day
#.default says that everything else should be left the same but +5 

mutate(election_day = as.numeric(election_day))
# as.numeric transforms the column into a numeric column 

clean_names()
#makes column names to lower case and replaces spaces with column dates 

clean_names() |> 
  rename(linenumber = x1_linenumber)
# renames the old column into the new one 
# ! new name comes first, then comes old name 

get_dupes()
# identifies two excact same columns 

problems(texas_precinct_20)
# identifies problems 

texas_precinct_20 <- texas_precinct_20 |>
  mutate(election_day = case_when(
    election_day == '5+' ~ '49',
    .default = election_day
  ))
  
#default leaves everything else as it is but the 5+
 
 cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount)) |> 
  distinct()
#gets rid of the dupes 
 
 cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount)) |> 
  distinct() |>
  mutate(zip = str_sub(zip, start=1L, end=5L)) |>
  mutate(city = str_to_title(city)) |>
  mutate(city = case_when(
    city == "Conowing" ~ "Conowingo",
    TRUE ~ city
    ))
    
install.packages('refinr')
# installs refinr package 

dc_food |>
  group_by(recipient_name) |>
  summarise(
    count=n()
  ) |>
  filter(str_detect(recipient_name, '^w|^W')) |> 
  arrange(recipient_name)
  
# The filter function in the codeblock below says: look in the city column, and pluck out any value that starts with (the "^" symbol means "starts with") a lowercase "w" OR (the vertical "|", called a pipe, means OR) an uppercase "W". 

cleaned_dc_food <- dc_food |>
  mutate(recipient_clean=key_collision_merge(recipient_name)) |>
  select(recipient_name, recipient_clean, everything())
  
# Key_collision_merge tries to find a standardized version of a certain value 

Open Refine 
# Tool by Google that allows standarizing values 

bind_rows()
# binds rows together 

left_join
# joins data from the left

right_join 
# joins data from the right
# important: order matters 

inner_join

join_by
# which column are we joining by
# !!! if join doesn't excatly match, it's not gonna match 

nrow()
# just show me the numbers of rows 

as.character
#turns values into characters

as.numeric
#turns values into numbers 

(across(ZCTA5N, as.character))
#across makes sure that every value gets changed 

