---
title: "Class Reference"
author: "Loretta Pulwer"
output:
  html_document:
    # theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```


library(tidyverse)
── Attaching core tidyverse packages ────────────────────── tidyverse 2.0.0 ──
✔ dplyr     1.1.4     ✔ readr     2.1.5
✔ forcats   1.0.0     ✔ stringr   1.5.1
✔ ggplot2   3.4.4     ✔ tibble    3.2.1
✔ lubridate 1.9.3     ✔ tidyr     1.3.1
✔ purrr     1.0.2     
── Conflicts ──────────────────────────────────────── tidyverse_conflicts() ──
✖ dplyr::filter() masks stats::filter()
✖ dplyr::lag()    masks stats::lag()
ℹ Use the conflicted package to force all conflicts to become errors
> #library loads packages. In this case, it's the tidyverse

> install.packages('tidyverse')
Error in install.packages : Updating loaded packages
> #this functon intalls packages 

> maryland_winred_contributions <- read_rds("maryland_winred.rds")
Error in readRDS(con, refhook = refhook) : cannot open the connection
In addition: Warning message:
In readRDS(con, refhook = refhook) :
  cannot open file 'maryland_winred.rds': No such file or directory
> #I don't have the file on my computer. But in theory, R studio is supposed to read to import the data from the given file 
