---
title: "lab_03"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

## Fix header issues

You'll need to fix one issue with the column names using janitor. Write a code block to do that, making sure that you save the results to the same dataframe.

```{r}
md_contribs_clean <- md_contribs |>
  clean_names()
```

## Answer questions

Q1: Which seven cities in Maryland had the largest *number* of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1: Bethesda (1177), Chevy Chase (704), Baltimore (690), Annapolis (490), Potomac (490), Silver Spring (405), Rockville (385)

```{r}
md_contribs_clean |>
  mutate(contributor_city = case_when(contributor_city == "Balrimore" ~ "Baltimore",
 contributor_city == "Silver Springs" ~ "Silver Spring", 
 contributor_city == "Rockwell" ~ "Rockville",
 contributor_city == "Potamic" ~ "Potomac",
 TRUE ~ contributor_city 
 ))
md_contribs |>
  group_by(contributor_city) |>
  summarise(
    count_contribs = n()
  ) |>
arrange(desc(count_contribs))
```

```{r}
distinct_variations_city <- md_contribs_clean |>
distinct(contributor_city)
#I used this function to check is there are any misspellings. 
```

Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). You will need to use mutate to ensure that you have a zip code column with the first five characters from the existing zip code column.

Where is that zip code and what kind of area is it, demographically? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: 20817 which applies to Bethesda which is one of the wealthiest cities in the country. 
SHEILA CHERFILUS MCCORMICK FOR CONGRESS, INC and YVETTE4CONGRESS

As far as I can see the top 5 candidates, It seems like none of them are candidates for Maryland (I googled where they are)

```{r}
md_contribs_clean |> 
 mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L)) |>
group_by(contributor_zip) |>
  summarise(
   count_contribs = n(),
  contribution_receipt_amount = sum(contribution_receipt_amount)
 ) |>
  arrange(desc(contribution_receipt_amount)) 
```

```{r}
md_contribs_clean |> 
  filter(contributor_zip == 20817) |>
group_by(committee_name) |>
  summarise(
 count_contribs = n(),
  contribution_receipt_amount = sum(contribution_receipt_amount)
 ) |>
arrange(desc(contribution_receipt_amount))

```

Q3: Let's look at top individual donors. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. What does the original data say about the top two donors - who are they and what do they do? Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?

A3: Top two contributors are Melanie Jacobsen and Mark P. Jacobsen from Bethesda. Melanie Jacobsen works fo Banking Services and so does Mark P. Jacobsen. Moreover, Mark P. Jacobsen is the CEO of IntraFi. 
Something that stood out when I glimpsed at the top 10 donors was that there is another Mark Jacobsen from Bethesda. So we have a duplicate here, meaning that Mr. Jacobsen actually contributed much more. Same applies to Jeffrey Drezner.

```{r}
top_donors <- md_contribs_clean |> 
  group_by(contributor_name, contributor_city) |> 
  summarise(total_contribution = sum(contribution_receipt_amount, na.rm = TRUE), .groups = 'drop') |> 
   arrange(desc(total_contribution))
#DISCLAIMER: I used Chatgpt after I typed the following code in: top_donors <- md_contribs_clean |> group_by(contributor_name, contributor_city) |>  summarise(total_contribution = sum(contribution_receipt_amount, na.rm = TRUE)) |> arrange (desc(total_contribution)). It gave me the following error message: "summarise() has grouped output by 'contributor_name'. You can override using the .groups argument." Chatgpt then added the .groups = 'drop'. I don't understand what I was missing. What does this do?

```
